---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data

```{r echo=TRUE}
unzip<-unzip("activity.zip",unzip="internal")
data<-read.csv(unzip, colClasses = c("integer", "Date", "factor"))
data$date<-as.Date(data$date, format = "%Y-%m-%d", usetz = FALSE)
```

## What is mean total number of steps taken per day?

Here's a histogram, representing the total number of steps taken per day.
``` {r echo = TRUE}
ag<-aggregate(steps ~ date, data = data, sum)
library(ggplot2)
qplot(steps, data = ag)
mean_steps<-mean(ag$steps)
median_steps<-median(ag$steps)
```

The mean of the total number of steps taken per day is `r mean_steps`. The median of the total number of steps taken per day is `r median_steps`.

## What is the average daily activity pattern?

``` {r echo = TRUE}
daily_pat<-aggregate(steps ~ interval, data = data, mean)
library(ggplot2)
g<-ggplot(daily_pat, aes(interval, steps))
g + geom_point(type = "l")
max_steps<-max(daily_pat$steps)
max_r<-which(daily_pat$steps == max_steps)
max_int<-daily_pat[max_r,1]
```

The `r max_int` 5-min interval on average across all the days in the dataset contains the maximum number of steps!

## Imputing missing values

```{r echo = TRUE}
z<-na.omit(data)
nNA<-nrow(data) - nrow(z)
sum(is.na(data$steps))
sum(is.na(data$interval))
sum(is.na(data$date))
```

The total number of observations with missing values in the dataset is `r nNA`.
We can see that only steps column has NAs
I will substitute them with mean data for every interval.

``` {r echo = TRUE}
NAs<-which(is.na(data$steps) == TRUE)

data_new<-data

for (i in 1:2304) {
        n<-NAs[i]
        data_new[n, 1] <- daily_pat[which(daily_pat$int == data_new[n, 3]),2]
}

ag_new<-aggregate(steps ~ date, data = data_new, sum)
library(ggplot2)
qplot(steps, data = ag_new)
mean_steps_n<-mean(ag_new$steps)
median_steps_n<-median(ag_new$steps)

```

The mean of the total number of steps taken per day is `r mean_steps_n`. The median of the total number of steps taken per day is `r median_steps_n`. Imputing the NAs have caused the median steps data to become noninteger. The total number of everyday steps have of course incresased!


## Are there differences in activity patterns between weekdays and weekends?

```{r echo = TRUE}
data_new$weekday<-weekdays(data_new$date)
data_new$day<-rep(NA)
for (i in 1:nrow(data_new)) {
        if (data_new[i,4] == "Saturday")
        {
                data_new[i,5] <- c("weekend")
        
        }
        else if (data_new[i,4] == "Sunday")
        {
                data_new[i,5] <- c("weekend")
        
        }
        
        else {
                data_new[i,5] <- c("weekday")
        }
} 

daily_pat_n<-aggregate(steps ~ interval + day, data = data_new, mean)
library(ggplot2)
g<-ggplot(daily_pat_n, aes(interval, steps))
g + geom_point(type = "l") + facet_wrap(~ day, nrow = 2, ncol = 1)
```

